The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-07-05 analysis/paper/paper.qmd
Add wip topic modeling script

## Topic model

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017]. We will use the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.

```{r topic-model-lda}
emu_theses_topics <- get_topics(emu_theses_words)

emu_theses_topics_terms <- tidy(emu_theses_topics, matrix = "beta")
emu_theses_topics_documents <- tidy(emu_theses_topics, matrix = "gamma")

# Prepare document-term matrix as input for LDA model
emu_dtm <- convert_to_dtm(emu_theses_words, "title", "word")

# Create LDA model
emu_lda <- LDA(emu_dtm, k = 2, control = list(seed = 2023))

@@

```{r topic-model-beta}

>>>>

### Topic model

Similar to clustering on numeric data, topic modeling is a method of unsupervised classification of topics found in a group of documents [@silge2017]. We used the Latent Dirichlet Allocation (LDA) algorithm for topic modeling. In LDA, each document is a mixture of topics and each topic is a mixture of words. The probability of a term being part of a topic is given by the $\beta$ (beta) statistic, while the probability of a document being part of a topic is given by the $\gamma$ (gamma) statistic.

The number of topics was empirically determined with the `ldatuning` package [@???].

```{r topic-model-lda}
#| dependson: short-stop-words

# Prepare document-term matrix as input for LDA model
# Reduce the size of the DTM by excluding words with a frequency lower than 5
emu_dtm <- convert_to_dtm(emu_theses_words, 
                          title_col = "title", 
                          word_col = "word", 
                          min_term_freq = 5)

# Fit LDA model with K clusters
# # Determine K value empirically
# K = get_number_of_topics(emu_dtm)
# # This returns K = 23
# TODO run the model with K = 23 and compare with results using qualitative value

# Determine K value qualitatively
# K = 5 corresponds to the 2 TUD semesters and the 3 exchange semesters combined 
emu_lda <- LDA(emu_dtm, k = 5, method="Gibbs",
               control = list(seed = 2023, iter = 500, verbose = 25))

# Extract beta and theta statistics from LDA model
beta <- posterior(emu_lda)$terms
theta <- posterior(emu_lda)$topics

# Preview 10 terms from each topic
terms(emu_lda, 10)

# Add topic pseudo-names based on the top 5 words in each topic
topic_names <- name_topics(beta)

@@

```{r topic-model-lda}

# What is the dominant topic in each document?
N = 20
topic_prop_examples <- theta[1:N,]
colnames(topic_prop_examples) <- topic_names
vizDataFrame <- melt(cbind(data.frame(topic_prop_examples), 
                           document = factor(rownames(topic_prop_examples))), 
                     variable.name = "topic", id.vars = "document")  
ggplot(data = vizDataFrame, aes(topic, value, fill = document), ylab = "proportion") + 
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip() +
  facet_wrap(~ document, ncol = 5) +
  theme(legend.position = "none")
```

```{r}

```{r}

# What are the most probable topics in the entire collection?
topicProportions <- colSums(theta) / nDocs(emu_dtm)  # mean probabilities over all paragraphs
names(topicProportions) <- topic_names     # assign the topic names we created before
sort(topicProportions, decreasing = TRUE) # show summed proportions in decreased order

soP <- sort(topicProportions, decreasing = TRUE)
paste(round(soP, 5), ":", names(soP))

# How many documents are with the highest probability in each topic?
K = 5
countsOfPrimaryTopics <- rep(0, K)
names(countsOfPrimaryTopics) <- topic_names
for (i in 1:nDocs(emu_dtm)) {
  topicsPerDoc <- theta[i, ] # select topic distribution for document i
  # get first element position from ordered list
  primaryTopic <- order(topicsPerDoc, decreasing = TRUE)[1] 
  countsOfPrimaryTopics[primaryTopic] <- countsOfPrimaryTopics[primaryTopic] + 1
}
sort(countsOfPrimaryTopics, decreasing = TRUE)

so <- sort(countsOfPrimaryTopics, decreasing = TRUE)
paste(so, ":", names(so))
```

```{r}
topicToFilter <- 6  # you can set this manually ...
# ... or have it selected by a term in the topic name (e.g. 'children')
topicToFilter <- grep('network', topicNames)[1] 
topicThreshold <- 0.2
selectedDocumentIndexes <- which(theta[, topicToFilter] >= topicThreshold)
filteredCorpus <- corpus[selectedDocumentIndexes]
# show length of filtered corpus
filteredCorpus
```

```{r}

N = 82
topic_prop_per_year <- theta[1:N,]
colnames(topic_prop_per_year) <- topic_names
vizDataFrame <- melt(cbind(data.frame(topic_prop_per_year), 
                           document = factor(rownames(topic_prop_per_year))), 
                     variable.name = "topic", id.vars = "document")

# plot topic proportions per decade as bar plot
vizDataFrameWithYear <- vizDataFrame |> 
  left_join(select(emu_theses, title, graduation_year), by = c("document" = "title")) |> 
  ggplot(aes(x=factor(graduation_year), y=value, fill = topic)) + 
  geom_bar(stat = "identity", position = "fill") +
  scale_fill_viridis_d(name = "Topics") +
  xlab("Graduation year") +
  ylab("Proportion of topics") +
  labs(title = "Proportion of summed topic probabilities over the years of the EMU program")

```{r lda-vis}

topicmodels_json_ldavis <- function(fitted, corpus, doc_term) {

  phi <- topicmodels::posterior(fitted)$terms |> as.matrix
  theta <- topicmodels::posterior(fitted)$topics |> as.matrix
  vocab <- colnames(phi)
  doc_length <- vector()

  for (i in 1:length(corpus)) {
    temp <- paste(corpus[[i]]$content, collapse = " ")
    doc_length <- c(doc_length, stringi::stri_count(temp, regex = '\\S+'))
  }

  temp_frequency <- doc_term |> as.matrix
  freq_matrix <- data.frame(ST = colnames(temp_frequency),
                            Freq = colSums(temp_frequency))
  rm(temp_frequency)

  json <- LDAvis::createJSON(phi = phi,
                             theta = theta,
                             doc.length = doc.length,
                             vocab = vocab,
                             term.frequency = freq_matrix$Freq)
}

LDAvis::serVis(emu_lda_json, out.dir = "vis", open.browser = FALSE)

```{r topic-model-beta}
#| dependson: topic-model-lda
